### 2021-11-13 NSF2021/Description/specificaim2.tex

When solving for the density functions $\eta_0$, $\eta_1$, and $\sigma$,
a carefully chosen IE formulation requires a mesh-independent number of
GMRES iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required
CPU time is proportional to the cost of a matrix-vector multiplication
that can be performed in optimal or near-optimal time with the fast
multipole method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2,
fmm3, fmm4, fmm6, fmm7, fmm8}. PI BQ is experienced with applying FMMs
for the Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the
screened Laplace equation~\cite{kro-qua2011, qua2011}. Furthermore, to
reduce the number of GMRES iterations, preconditioning strategies such
as the inverse Fast Multipole Method (IFMM)~\cite{cou-pou-dar2017} and
domain decomposition~\cite{che-bir2021}. PI BQ used the IFMM to
precondition Stokes equations~\cite{qua-cou-dar2018}. A suite of other
preconditioners including sparse approximate inverses~\cite{che2000} and
multigrid~\cite{hem-sch1981, sch1982} will be investigated.

@@

A numerical issue when simulating the self-assembly of amphiphilic
particles is avoiding particle collision. The hydrophobic attraction
potential drives the amphiphilic particles towards one another and this
leads to physical contact in finite time. Such particle collisions in a
dense, rigid-body suspension are a great challenge and can be a
bottleneck in large-scale simulations. We propose to remedy this
computational challenge by developing novel high-order adaptive time
stepping methods. PI BQ developed a high-order adaptive time stepping
method for hydrodynamic suspensions \cite{qua-bir2016}, and it has
served as a robust method to simulate processes including mixing and
adhesion in suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}.
Following this previous work, the proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order, single-step method to achieve high-order accuracy.
To estimate the error, at each time step, we will compute the total
force and torque of the system which is computationally cheap and
physically zero. PI BQ's experience is that error estimates based on
physical constraints such as force- and torque-free conditions
appropriately determine how the time step should be adjusted so that the
dynamics are resolved without the computational expense of techniques
such as embedded Runge-Kutta methods, step-doubling, and Richardson
extrapolation.

%%
%\begin{wrapfigure}[13]{r}{0.30\textwidth}
%\centerline{\includegraphics[width=0.29\textwidth]{figures/BIError.pdf}}
%  \vspace{-8pt}
%\caption{
%\label{fig:bierror}  
%\footnotesize The false color map shows how numerical quadrature of
%  layer potentials loses accuracy near the particle boundaries.  The
%  color bar is for $\log_{10}.$}
%\end{wrapfigure}
%%
%Our previous work avoided particle contact through Lennard-Jones body
%forces~\cite{Fu2018_SIAM}. Such steep, short-range, steric interactions 
%introduce numerical stiffness and limit the time step size. To
%remove this stiffness, we propose a geometry-based contact
%method~\cite{har-pon-sor-zor2011}. This method has been applied to
%vesicle suspensions~\cite{lu-rah-zor2017} and to rigid body suspensions in
%two dimensions~\cite{bys-sha-qua2020} and three
%dimensions~\cite{Yan2019}. These methods determine repulsion force by
%solving a non-linear complementarity problem with a geometric constraint
%that the configuration is non-overlapping. In this manner, contact is
%avoided without the excessively small time steps required by stiff
%Lennard-Jones forces.
%
%%-------------------------------------------------------------------------------
%\subsubsection{Novel reciprocal identities}
%\label{subsec:reciprocal}
%The force and torque formulas~\eqref{forceandtorque} use the
%hydrophobic stress along the particle boundaries. Unfortunately,
%layer potentials for $\nabla u$ involve hyper-singular integrals and
%this introduces large amounts of quadrature error on the particle
%boundaries (Figure~\ref{fig:bierror}). Therefore, it is useful to devise
%reciprocal identities for the force and torque on body $i$ that do not
%require integration along body $i$. One identity, that we have proved,
%is
%\begin{align}
%    \label{eq:reciprocal}
%{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
%{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
%  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
%  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
%\end{align}
%for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
%equation when only the contribution from particle $i$ is considered and
%$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j \mathbf{I} +
%\rho\left((\nabla u_i \cdot \nabla u_j) \mathbf{I} - 2 \nabla u_i
%\otimes \nabla u_j\right)$. Equation~\eqref{eq:reciprocal} does not
%require knowledge of $\nabla u$ on particle boundary $i$, and this is
%enormously beneficial for calculating the hydrophobic force and torque.
%
%% ----------------------------------------------------------------------
%\subsubsection{Fluctuating hydrodynamics}
%\label{subsec:fluctuating}
%Once algorithms that address quadrature, adaptive time stepping, and
%contact are implemented, we will embark on incorporating fluctuating
%hydrodynamics that are especially important at scales of amphiphilic
%particles. Bao et al.~\cite{Bao17,Bao18} developed two-dimensional BIE
%formulations for Brownian rigid bodies. To satisfy the covariance of the
%rigid body velocities, they use an ill-conditioned, first-kind integral
%equation method. The proposed work will consider a layer potential
%formulation that has much better conditioning properties. This strategy
%was attempted in~\cite{Bao18}, but they claim that a necessary
%regularization of the covariance matrix leads to a drastic loss of
%accuracy in numerical fluctuation-dissipation balance. However, only one
%regularization strategy was considered, and as PI BQ has shown in
%previous work, the regularization choice significantly affects the
%resulting physics~\cite{ong-chr-qua2017}. Therefore, different
%regularization strategies will be investigated with the goal of finding
%well-conditioned BIE formulations that control the accuracy of the
%fluctuation-dissipation balance.

>>>>

When solving for the density functions $\eta_0$, $\eta_1$, and $\sigma$,
a carefully chosen IE formulation requires a mesh-independent number of
GMRES iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required
CPU time is proportional to the cost of a matrix-vector multiplication
that can be performed in optimal or near-optimal time with the fast
multipole method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2,
fmm3, fmm4, fmm6, fmm7, fmm8}. PI Quaife is experienced with applying
FMMs for the Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the
screened Laplace equation~\cite{kro-qua2011, qua2011}. Furthermore, to
reduce the number of GMRES iterations, preconditioning strategies such
as the inverse Fast Multipole Method (IFMM)~\cite{cou-pou-dar2017} and
domain decomposition~\cite{che-bir2021}. PI Quaife used the IFMM to
precondition Stokes equations~\cite{qua-cou-dar2018}. A suite of other
preconditioners including sparse approximate inverses~\cite{che2000} and
multigrid~\cite{hem-sch1981, sch1982} will be investigated.

@@

A numerical issue when simulating the self-assembly of amphiphilic
particles is avoiding particle collision. The hydrophobic attraction
potential drives the amphiphilic particles towards one another and this
leads to physical contact in finite time. Such particle collisions in a
dense, rigid-body suspension are a great challenge and can be a
bottleneck in large-scale simulations. We propose to remedy this
computational challenge by developing novel high-order adaptive time
stepping methods. PI Quaife developed a high-order adaptive time
stepping method for hydrodynamic suspensions \cite{qua-bir2016}, and it
has served as a robust method to simulate processes including mixing and
adhesion in suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}.
Following this previous work, the proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order, single-step method to achieve high-order accuracy.
To estimate the error, at each time step, we will compute the total
force and torque of the system which is computationally cheap and
physically zero. PI Quaife's experience is that error estimates based on
physical constraints such as force- and torque-free conditions
appropriately determine how the time step should be adjusted so that the
dynamics are resolved without the computational expense of techniques
such as embedded Runge-Kutta methods, step-doubling, and Richardson
extrapolation.

### 2021-11-13 NSF2021/Description/specificaim3.tex

With these results, the PIs will be ready to make quantitative
connections between a Janus particle membrane and a lipid bilayer
membrane in an electrolytic solution. In recent
experiments~\cite{FaizEtAl2019_SoftMatt}, membrane bending rigidity was
determined as a function of lipid composition from 0 to 100 mol $\%$ of
charged lipids using flicker spectroscopy of the shape fluctuations of a
giant unilamellar vesicle (GUV). Membrane bending rigidity increases
with increasing lipid surface charge, but decreases with increasing salt
concentration in the bulk solution due to the screening of the lipid
surface charge. This result agrees with several theoretical
models~\cite{Kralchevsky1996_JCIS, May1996_JChemPhys,
LoubetEtAl2013_PRE} that also assume the quadratic form of the elastic
energy density in the presence of surface charge and bulk
charge~\cite{DuplantierGoldstein1990_PRL, Winterhalter1992_JPC}. As the
electrostatic interaction is non-local in nature, we expect that the
controversy of the HK elastic energy form (see
\S\ref{subsec:specific_aim_1}) would worsen in the presence of
electrostatic interactions and electrokinetics. The PIs will extend the
approaches in \S\ref{subsec:specific_aim_1} to charged lipids to
calculate the bending moduli and compare against the experimental
results in~\cite{FaizEtAl2019_SoftMatt}. We propose to incorporate an
explicit surface charge on each particle boundary $P_i$ and compute the
electrostatic potential as a functional of particle configuration.  By
adding the electrostatic force to the hydrophobic attraction force
between particles, we will assess the dependence of elastic moduli on
electric charges using the methods described in
\S\ref{subsec:specific_aim_1}. Results from these proposed calculations
will provide further comparisons and validations of the HAP model
against the continuum mechanics.

Once the effects of lipid charges on the moduli are verified, the PIs
propose to examine how to use an external electric field to control the
amphiphilic self-assembly in solvent. PI Young has a track record of
working on electrohydrodynamics of an elastic, inextensible membrane
using both asymptotic analysis~\cite{Nganguia2013_PRE, Young2014_JFM,
Young2015_PoF} and numerical simulations~\cite{Nganguia2015_CiCP} and
will work with both PIs Ryham and Quaife to extend the HAP model to
study the electromechanical effects on the assembly of amphiphile.
Results from this research will yield a quantitative understanding of
how to utilize an electric field to achieve optimal control of assembly
of amphiphiles in solvent.

>>>>

With these results, the PIs will be ready to make quantitative
connections between a Janus particle membrane and a lipid bilayer
membrane in an electrolytic solution. In recent
experiments~\cite{FaizEtAl2019_SoftMatt}, membrane bending rigidity was
determined as a function of lipid composition from 0 to 100 mol $\%$ of
charged lipids using flicker spectroscopy of the shape fluctuations of a
giant unilamellar vesicle (GUV). Membrane bending rigidity increases
with increasing lipid surface charge, but decreases with increasing salt
concentration in the bulk solution due to the screening of the lipid
surface charge. This result agrees with several theoretical
models~\cite{Kralchevsky1996_JCIS, May1996_JChemPhys,
LoubetEtAl2013_PRE} that also assume the quadratic form of the elastic
energy density in the presence of surface charge and bulk
charge~\cite{DuplantierGoldstein1990_PRL, Winterhalter1992_JPC}. As the
electrostatic interaction is non-local in nature, we expect that the
controversy of the HK elastic energy form (see
\S\ref{subsec:specific_aim_1}) would worsen in the presence of
electrostatic interactions and electrokinetics. The PIs will extend the
approaches in \cite{Fu2018_SIAM} to charged lipids to
calculate the bending moduli and compare against the experimental
results in~\cite{FaizEtAl2019_SoftMatt}. We propose to incorporate an
explicit surface charge on each particle boundary $\Gamma_i$ and compute the
electrostatic potential as a functional of particle configuration.  By
adding the electrostatic force to the hydrophobic attraction force
between particles, we will assess the dependence of elastic moduli on
electric charges using the methods described in \cite{Fu2018_SIAM}. Results from these proposed calculations
will provide further comparisons and validations of the HAP model
against the continuum mechanics.

Once the effects of lipid charges on the moduli are verified, the PIs
propose to examine how to use an external electric field to control the
amphiphilic self-assembly in solvent. PI Young has a track record of
working on electrohydrodynamics of an elastic, inextensible membrane
using both asymptotic analysis~\cite{Nganguia2013_PRE, Young2014_JFM,
Young2015_PoF} and numerical simulations~\cite{Nganguia2015_CiCP} and
will work with both PIs Ryham and Quaife to extend the HAP model to
study the electromechanical effects on the assembly of colloidal amphiphiles.
Results from this research will yield a quantitative understanding of
how to utilize an electric field to achieve optimal control of assembly
of amphiphiles in solvent.
